# Exercise 1: Orange juice dataset

The `OJ` data from the `ISLR` package contains 1070 purchases where the customer either purchased Citrus Hill or Minute Maid Orange Juice. A number of characteristics of the customer and product are recorded.

* `Purchase`: A factor with levels CH and MM indicating whether the customer purchased Citrus Hill or Minute Maid Orange Juice (i.e. response/class)
* `WeekofPurchase`: Week of purchase
* `StoreID`: Store ID
* `PriceCH`: Price charged for CH
* `PriceMM`: Price charged for MM
* `DiscCH`: Discount offered for CH
* `DiscMM`: Discount offered for MM
* `SpecialCH`: Indicator of special on CH
* `SpecialMM`: Indicator of special on MM
* `LoyalCH`: Customer brand loyalty for CH
* `SalePriceMM`:Sale price for MM
* `SalePriceCH`: Sale price for CH
* `PriceDiff`: Sale price of MM less sale price of CH
* `Store7`: A factor with levels No and Yes indicating whether the sale is at Store 7
* `PctDiscMM`: Percentage discount for MM
* `PctDiscCH`: Percentage discount for CH
* `ListPriceDiff`: List price of MM less list price of CH
* `STORE`: Which of 5 possible stores the sale occured at

Source: Stine, Robert A., Foster, Dean P., Waterman, Richard P. Business Analysis Using Regression (1998). Published by Springer.

To load the data, use
```{r}
library(ISLR)
data(OJ)
```

## Exploratory data analysis

In `R`, categorical variables are sometimes recorded as numerical variables, which is inappropriate for subsequent analysis. To address this, `as.factor()` should be used to convert these variables back into categorical variables. 

**Task**

(a) Use `str`, `summary` and/or `skim` to inspect the data. Looking at the variable description above, is there any categorical variable encoded incorrectly as numerical variable? If so, correct them by using `as.factor()`. 

Excluding the response variable `Purchase`, how many variables are numerical variables and how many variables are categorical variables?

* The number of numerical variables = `r fitb(12)`.
* The number of numerical variables = `r fitb(5)`.

```{r, webex.hide="Solution"}
library(skimr)
str(OJ)
skim(OJ)
OJ$StoreID <- as.factor(OJ$StoreID)
OJ$SpecialCH <- as.factor(OJ$SpecialCH)
OJ$SpecialMM <- as.factor(OJ$SpecialMM)
OJ$STORE <- as.factor(OJ$STORE)
str(OJ)
skim(OJ)
```

(b) 
